Sustainable computational science:\\ the ReScience initiative

### 2017-11-07 replies_to_reviewers.md

# Replies to the reviewers' comments

## In reply to reviewer 1

## In reply to reviewer 2

We thank the reviewer for his enthusiastic comments, and we would be
happy to welcome him as a contributor to ReScience.

We have corrected the name of Jonathan Buckheit.

### 2017-11-08 replies_to_reviewers.md

# Replies to the reviewers' comments

## In reply to reviewer 1

We thank the reviewer for the in-depth examination of our work and for his useful suggestions for the future evolution of ReScience.

Concerning what the reviewer calls a defeatist attitude towards computational reproducibility, that is not at all how we see our approach. The manuscript focuses on reporting our experiences with ReScience. Concerning reproducibility, we note that having it verified by reviewers is not sufficient. We are not aware of any practical experience demonstrating that some other approach is efficient, which is why we consider the reproducibility of software environments an unsolved problem. Many ideas are being explored at the moment, including containerization as the reviewer mentions. The effectiveness of these ideas cannot be judged yet. Nobody knows, for example, if five years from now, the then-current version Docker will still be able to handle today's container images. We see our attitude not as defeatist, but as prudent. Many of the authors are contributing to the development and evaluation of new ideas, and we will certainly try some of them out in the context of ReScience.

We have reworded some of our statements to make clearer that our doubts about reproducibility apply only to today's state of the start. We have also added a new paragraph on the reproducibility of software environments to the Outlook section.

### 2017-11-08 draft.tex
Changes in reply to reviewer 2
Reproducibility of software environments
In reply to reviwer #1

\twocolumn[
  \maketitle
  \vspace{5mm}
  \noindent
  \textbf{Abstract.}
  %
  Computer science offers a large set of tools for prototyping, writing,
  running, testing, validating, sharing and reproducing results, however
  computational science lags behind.  In the best case, authors may provide
  their source code as a compressed archive and they may feel confident their
  research is reproducible. But this is not exactly true. James Buckheit and
  David Donoho proposed more than two decades ago that an article about
  computational results is advertising, not scholarship. The actual scholarship
  is the full software environment, code, and data that produced the
  result. This implies new workflows, in particular in peer-reviews. Existing
  journals have been slow to adapt: source codes are rarely requested, hardly
  ever actually executed to check that they produce the results advertised in
  the article.
  %
  ReScience is a peer-reviewed journal that targets computational research and
  encourages the explicit replication of already published research, promoting
  new and open-source implementations in order to ensure that the original
  research can be replicated from its description. To achieve this goal, the
  whole publishing chain is radically different from other traditional
  scientific journals. ReScience resides on GitHub where each new
  implementation of a computational study is made available together with
  comments, explanations, and software tests.\\

@@

While some of the reasons for non-reproducibility are specific to each
scientific domain, our experience has shown that there are also some
common issues that can be identified.  Missing code and/or data,
undocumented dependencies, and inaccurate or imprecise description
appear to be characteristic of much non-reproducible work. Moreover,
these problems are not always easy to detect even for attentive
reviewers, as we discovered when some articles published in ReScience
turned out to be difficult to reproduce for someone else for exactly
the reasons listed above. ReScience reviewers are scientists working
in the same domain as the submitting authors, because familiarity with
the field is a condition for judging if a replication is
successful. But this also means that our reviewers share a significant
common background with the authors, and that background often includes the
software packages and programming languages adopted by their
community.  In particular, if both authors and reviewers have
essential libraries of their community installed on their computers,
they may not notice that these libraries are actually dependencies of
the submitted code.  While solutions to this problem evidently exist
(ReScience could, for example, request that authors make their
software work on a standard computational environment supplied in the
form of a virtual machine), they represent an additional effort to
authors and therefore discourage them from submitting replication work to
ReScience.  Our goal is to develop a set of recommendations for
authors that represent a workable compromise between reproducibility
and implementation effort.\\

Another issue with reproducibility is that with today's computing
technology, long-term reproducibility can only be achieved by imposing
drastic constraints on languages and libraries that are not compatible
with the requirements of research computing. This problem is nicely
illustrated by \citet{Mesnard:2016} whose authors report trying to
reproduce their own work performed two years earlier. Even though
Barba's group is committed to reproducible research practices, they
did not escape the many problems one can face when trying to re-run a
piece of code. As a consequence, the code newly produced for ReScience
will likely cease to be functional at some point in the future.
Therefore, the long-term value of a ReScience publication is not the
actual code but the accompanying article. The combination of the
original article and the replication article provide a complete and
consistent description of the original work, as evidenced by the fact
that replication was possible. Even 5, 10, or 20 years later, a
competent scientist should be able to replicate the work again thanks
to these two articles. Of course, the new code can also help, but the
true long-term value of a replication is the accompanying article.\\

@@

>>>>

\twocolumn[
  \maketitle
  \vspace{5mm}
  \noindent
  \textbf{Abstract.}
  %
  Computer science offers a large set of tools for prototyping, writing,
  running, testing, validating, sharing and reproducing results, however
  computational science lags behind.  In the best case, authors may provide
  their source code as a compressed archive and they may feel confident their
  research is reproducible. But this is not exactly true. Jonathan Buckheit and
  David Donoho proposed more than two decades ago that an article about
  computational results is advertising, not scholarship. The actual scholarship
  is the full software environment, code, and data that produced the
  result. This implies new workflows, in particular in peer-reviews. Existing
  journals have been slow to adapt: source codes are rarely requested, hardly
  ever actually executed to check that they produce the results advertised in
  the article.
  %
  ReScience is a peer-reviewed journal that targets computational research and
  encourages the explicit replication of already published research, promoting
  new and open-source implementations in order to ensure that the original
  research can be replicated from its description. To achieve this goal, the
  whole publishing chain is radically different from other traditional
  scientific journals. ReScience resides on GitHub where each new
  implementation of a computational study is made available together with
  comments, explanations, and software tests.\\

@@

While some of the reasons for non-reproducibility are specific to each
scientific domain, our experience has shown that there are also some
common issues that can be identified.  Missing code and/or data,
undocumented dependencies, and inaccurate or imprecise description
appear to be characteristic of much non-reproducible work. Moreover,
these problems are not always easy to detect even for attentive
reviewers, as we discovered when some articles published in ReScience
turned out to be difficult to reproduce for someone else for exactly
the reasons listed above. ReScience reviewers are scientists working
in the same domain as the submitting authors, because familiarity with
the field is a condition for judging if a replication is
successful. But this also means that our reviewers share a significant
common background with the authors, and that background often includes the
software packages and programming languages adopted by their
community.  In particular, if both authors and reviewers have
essential libraries of their community installed on their computers,
they may not notice that these libraries are actually dependencies of
the submitted code.  While solutions to this problem evidently exist
(ReScience could, for example, request that authors make their
software work on a standard computational environment supplied in the
form of a virtual machine), they represent an additional effort to
authors and therefore discourage them from submitting replication work to
ReScience.  Our goal is to develop a set of recommendations for
authors that represent a workable compromise between reproducibility
and implementation effort. These recommendations will evolve over time,
and we hope that with improving technology we will ultimately reach full
reproducibility over a few decades.\\

Another issue with reproducibility is that with today's computing
technology, long-term reproducibility can only be achieved by imposing
drastic constraints on languages and libraries that are not compatible
with the requirements of research computing. This problem is nicely
illustrated by \citet{Mesnard:2016} whose authors report trying to
reproduce their own work performed two years earlier. Even though
Barba's group is committed to reproducible research practices, they
did not escape the many problems one can face when trying to re-run a
piece of code. As a consequence, code that is written for ReScience
today will likely cease to be functional at some point in the future.
The long-term value of a ReScience publication lies not just in the
actual code but also in the accompanying article. The combination of the
original article and the replication article provide a complete and
consistent description of the original work, as evidenced by the fact
that replication was possible. Even 5, 10, or 20 years later, a
competent scientist should be able to replicate the work again thanks
to these two articles. Of course, the new code can also help, but the
true long-term value of a replication is the accompanying article.\\

@@

% -> Reproducible software environments

A fundamental technical issue is the difficulty of archiving or
accurately describing the software environments in which computational
scientists perform their work. A publication should be accompanied by
both a human-readable description of this environment and an
executable binary form. The human-readable description allows an
inspection of the versions of all software packages that were used,
for example to check for the impact of bugs that become known only
after a study was published. The executable version enables other
scientists to re-run the analyses and inspect intermediate
results. Ideally, the human-readable description would permit
rebuilding the executable version, in the same way that software
source code permits rebuilding executable binaries. This approach is
pursued for example by the package manager Guix \citep{Courtes:2015}.
A more limited but still useful implementation of the same idea exists
in the form of the conda package manager
\citep{conda}. Containerization, as implemented e.g. by Docker
\citep{docker} is currently much discussed, but provides only the
executable version without a human-readable description. Moreover, the
long-term stability of the container file format remains to be
evaluated. History has shown that long-term stability in computing
technology is achieved only by conventions for which it is a design
priority, for example the Java Virtual Machine \citep{JVM}. Docker, on
the contrary, is developed as a deployment technology with no ambition
towards long-term archiving.
